---
title: "Data 607- Week 7 assignment"
author: "Adriana Medina"
date: "2024-03-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval=FALSE, include=FALSE}
install.packages("xml2")
```
# Introduction
For this assignment I have prepared three separate files in HTML, XML, and JSON formats, each containing the following information about my favorite books:

 + Title
 + Author
 + Published Year
 + Genre

Each of the files were loaded into Github and then into R.
<header>Load necessary libraries:</header></b>

```{r}
library(tidyverse)
library(jsonlite)
library(xml2)
library(rvest)
```


# Load HTML file into a data frame
```{r}
htmlurl <- "https://raw.githubusercontent.com/amedina613/Data607-Week-7-Assignment/main/books.html"
html_data <- read_html(htmlurl)

# The HTML file loaded as class "xml_document"
class(html_data)
```

<header>Extract HTML table and convert it to a data frame:</header></b>

```{r}
html_table <- html_data %>%
  html_table(fill = TRUE)
 
html_df <- as.data.frame(html_table)
```


# Load XML file into a data frame

```{r}
xmlurl <- ("https://raw.githubusercontent.com/amedina613/Data607-Week-7-Assignment/main/books.xml")

xml_data <- read_xml(xmlurl)
#The xml file is loaded as class "xml_document"
class(xml_data)
```

<header>Extract the information from the XML:</header></b>

```{r}
titles <- xml_text(xml_find_all(xml_data, ".//title"))
authors <- xml_text(xml_find_all(xml_data, ".//authors"))
published_years <- as.numeric(xml_text(xml_find_all(xml_data, ".//published_year")))
genres <- xml_text(xml_find_all(xml_data, ".//genre"))

```

<header>Create a data frame for the information:</header></b>

```{r}
xml_df <- data.frame(
  title = titles,
  authors = authors,
  published_year = published_years,
  genre = genres
)
```

# Load JSON file into a data frame

<header>Read JSON data from URL:</header></b>

```{r}
json_url <- "https://raw.githubusercontent.com/amedina613/Data607-Week-7-Assignment/main/books.json"

json_data <- fromJSON(json_url)
#The JSON file is loaded as class data.frame
class(json_data)
```
<header>Convert to data frame:</header></b>

```{r}
json_df <- as.data.frame(json_data)
```

<header>View the data frames:</header></b>

```{r}
print(html_df)
print(xml_df)
print(json_df)
```
# Conclusion
There are a couple of differences in the naming of the columns. In the HTML data frame, the published year column is named "Published.Year" while in the other two they're named "published_year." The differences in column names could easily be fixed.

<header>Standardize column names:</header></b>
```{r}
names(html_df) <- c("title", "authors", "published_year", "genre")
names(xml_df) <- c("title", "authors", "published_year", "genre")
names(json_df) <- c("title", "authors", "published_year", "genre")
```

Another difference I noticed was the class of the imported data from each file. 
The HTML File(parsed with rvest):

 + "xml_document"
 
 The XML File(Parsed with xml2):
 
 + "xml_document"
 
 The JSON File (Parsed with jsonlite):
 
 + "data.frame"
